---
title: "Assignment2_Q4"
author: "Echo"
date: "12/02/2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(ggplot2)
library(ggpubr)

primer_type <- c(rep(1,3), rep(2,3), rep(3,3))
primer_type <- factor(rep(primer_type,2))
y_dipping <- c(4.0, 4.5, 4.3, 5.6, 4.9, 5.4, 3.8, 3.7, 4.0)
y_spraying <- c(5.4, 4.9, 5.6, 5.8, 6.1, 6.3, 5.5, 5.0, 5.0)
mydata <- data.frame(primertype = primer_type, application_methods = c(rep("dipping", 9), rep("spraying", 9)), adhesion_force = c(y_dipping, y_spraying))
```

1)

```{r}
plot1 <- ggplot(mydata, aes(x = primertype, y = adhesion_force)) + geom_boxplot()
plot2 <- ggplot(mydata, aes(x =application_methods, y = adhesion_force)) + geom_boxplot()

ggarrange(plot1, plot2, labels = c("A", "B"), ncol = 2)
```
The box plots are shown above. From the two boxplots, there seem to be difference in mean adhesion force between primer type groups (mean adhesion force in primer type 2 seem to be larger than other primer types). Also, there is difference in mean adhesion force in difference application methods. It looks like there is a smaller mean adhesion force with dipping than spraying.

2)

|Primer Type|Dipping|Spraying|Average for Primer Type|
|:---------:|:-----:|:------:|:---------------------:|
|1|4.0, 4.5, 4.3, $\bar{y}_{11}.=4.27$|5.4, 4.9, 5.6, $\bar{y}_{12}.=5.3$|$\bar{y}_{1..}=4.78$|
|2|5.6, 4.9, 5.4, $\bar{y}_{21}.=5.3$|5.8, 6.1, 6.3, $\bar{y}_{22}.=6.07$|$\bar{y}_{2..}=5.68$|
|3|3.8, 3.7, 4.0, $\bar{y}_{31}.=3.83$|5.5, 5.0, 5.0, $\bar{y}_{32}.=5.17$|$\bar{y}_{3..}=4.5$|
|Average for Application methods|$\bar{y}_{.1.}=4.47$|$\bar{y}_{.2.}=5.51$|$\bar{y}_{...}=4.99$|

Sum of Square of Total:
$$SST=\sum_{i=1}^{a}\sum_{j=1}^{b}\sum_{k=1}^{n}(\bar{y}_{ijk}-\bar{y}_{...})^2=\sum_{i=1}^{3}\sum_{j=1}^{2}\sum_{k=1}^{3}(\bar{y}_{ijk}-4.99)^2=10.7178$$

Sum of Square of Primer Type (Factor A): $$SS_A=bn\sum_{i=1}^{a}(\bar{y}_{i..}-\bar{y}_{...})^2=2\times3\sum_{i=1}^{3}(\bar{y}_{i..}-4.99)^2=4.5618$$

Sum of Square of Application Method (Factor B):
$$SS_B=an\sum_{j=1}^{b}(\bar{y}_{.j.}-\bar{y}_{...})^2=3\times3\sum_{j=1}^{2}(\bar{y}_{.j.}-4.99)^2=4.8672$$

Sum of Square of Interaction ($A\times B$):
$$SS_{AB}=n\sum_{i=1}^{a}\sum_{j=1}^{b}(\bar{y}_{ij.}-\bar{y}_{i..}-\bar{y}_{.j.}+\bar{y}_{...})^2=3\sum_{i=1}^{3}\sum_{j=1}^{2}(\bar{y}_{ij.}-\bar{y}_{i..}-\bar{y}_{.j.}+4.99)^2=0.2448$$

Sum of Square of Residual (RSS)
$$RSS=\sum_{i=1}^{a}\sum_{j=1}^{b}\sum_{k=1}^{n}(\bar{y}_{ijk}-\bar{y}_{ij.})^2=\sum_{i=1}^{3}\sum_{j=1}^{2}\sum_{k=1}^{3}(\bar{y}_{ijk}-\bar{y}_{ij.})^2=0.9868$$
The ANOVA table is shown below:

|Source of variation|SS|df|MS|
|:-----------------:|:--:|:--:|:--:|
|Primer Type|4.5618|2|2.2809|
|Application Method|4.8672|1|4.8671|
|Primer Type*Application Method|0.2448|2|0.1224|
|Residual|0.9868|12|0.0822|
|Total|10.7178|17||

3)

```{r}
interaction.plot(x.factor = mydata$primertype, trace.factor = mydata$application_methods, response = mydata$adhesion_force, fun = mean, col = c('blue', 'red'), lty = 1, lwd = 2, main = 'interaction plot', xlab = 'Primer Type', ylab = 'Adhesion Force', trace.label = 'Application method')
```
The interaction plot shows that lines for spraying and dipping seem to be parallel, and thus there is no interaction between Primer Type and Application Method.

4)

```{r}
anova1 <- aov(mydata$adhesion_force~mydata$primertype*mydata$application_methods)
summary(anova1)
```

The ANOVA table is shown above

5)

From ANOVA table, the test statistics for interaction term is F=1.466, with degree of freedom (2, 12). p-value is 0.269, which is greater than significant level $\alpha=0.05$. Thus, we fail to reject the null hypothesis and conclude that there is no significant interaction between Primer Type and Application Methods. 

6)
From ANOVA table, the test statistics for the effect of Application methods is F=59.703, with degree of freedom (1, 12). p-value is $5.36\times10^{-6}$, which is smaller than significant level $\alpha=0.05$. Thus, we reject the null hypothesis and conclude that application method affects the adhesion force. 
