---
title: "Assignment2_Q3"
author: "Echo"
date: "10/02/2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1)

```{r}
ya <- c(67, 62, 57, 63)
yb <- c(63, 59, 54, 59)
yc <- c(77, 71, 67, 74)
response <- c(ya, yb, yc)
treatment1 <- c(rep("A", 4), rep("B", 4), rep("C", 4))
treatment2 <- rep(c(1, 2, 3, 4), 3)
treatment2 <- factor(treatment2)
mydata <- data.frame(brand=treatment1, expertise=treatment2, y=response)
attach(mydata)
anova1 <- aov(y~brand+expertise)
summary(anova1)
```
The ANOVA table is shown above.

2)

The test statistics (F statistics) to is 441.8, with degree of freedom (2, 6). P-value is $3.07\times10^{-7}$, which is smaller than significant level $\alpha=0.05$. Thus, we reject the null hypothesis and conclude that there is differences exist between the effects of the keyboard brands on the mean number of words entered per minute.

3)

The test statistics (F statistics) to is 107.7, with degree of freedom (3, 6). P-value is $1.32\times10^{-5}$, which is smaller than significant level $\alpha=0.05$. Thus, we reject the null hypothesis and conclude that there is differences exist between the effects of the data entry specialists on the mean number of words entered per minute.

4)

```{r}
TukeyHSD(anova1, which = "brand")
detach(mydata)
```
From Tukeys simultaneous confidence intervals, Brand C maximize the mean words entered per minute because it the confidence interval for C-A and C-B are both above 0. 
